### 2017-03-15 doc/evaluation.tex
various late changes

\documentclass[12pt]{article}
\begin{document}
\section{Evaluation}
\subsection{Benchmarks}
\subsubsection{Parsing C Code}

\begin{itemize}
	\item Case studies from SPLVerifier (contributed to SVComp)
	\begin{itemize}
		\item email client
		\item elevator
		\item minepump
	\end{itemize}

	\item \emph{Stretch goal:} x86 Linux kernel files (about 7600 files) used in the evaluation of Typechef and SuperC
\end{itemize}

\subsubsection{C Code Typechecking}
On the same case studies used in C code parsing.
We need a C type checker written in a functional language though to lift.

\subsubsection{Simple LTS Algorithms}
\begin{itemize}
	\item State reachability
	\item State reachability with witness
\end{itemize}

\subsubsection{Model Transformations (graph rules)}
ICSE'14 paper

\subsection{Evaluation Metrics}
\subsubsection{Correctness}
Especially for the parsing case study, we can compare against Typechef and SuperC.

\subsubsection{Performance}
With respect to number of features and number of valid products.

Performance is evaluated for the baseline (black-box shallow lifting) and any optimizations that might be implemented. Possible optimizations include:
\begin{itemize}
	\item Using Binary Decision Diagrams (BDDs) instead of SAT solvers
	\item Deep-lifting of the List data structure for space optimization and data redundancy elimination
	\item Deep-lifting of other data structures
	\item Depp-lifting of algorithms (as needed, based on profiling results) 
\end{itemize}

\end{document}