### 2017-09-21 doc/taxonomy.tex

\documentclass{article}

%% macros
% reference macros
\newcommand{\Classen}{\cite{Classen:2010:MCL:1806799.1806850}}
\begin{document}

\title{A Taxonomy of Software Product Line Analyses}
\maketitle

\abstract
TODO

\section{Introduction}
TODO

\section{Definitions}

\section{Taxonomy}

\subsection{Model Checking}

\subsubsection{FTS}
A Featured-Transition System (FTS)\Classen is a modeling abstraction that extends a Labeled Transition System (LTS) to represent a family (product line) of models rather than a single model. 
\subsection{Data-Flow Analyses}
\subsection{Abstract Interpretation}
\subsection{Parsing}

\section{Discussion}
TODO

\section{Conclusion}
TODO

\bibliographystyle{ieeetr}
\bibliography{taxonomy} 
\end{document}