### 2017-09-22 doc/vamos18/taxonomy.tex
basic definitions

\documentclass{article}

%% macros
% reference macros
\newcommand{\Classen}{\cite{Classen:2010:MCL:1806799.1806850}}
\begin{document}

\title{A Taxonomy of Software Product Line Analyses}
\maketitle

\abstract
TODO

\section{Introduction}
TODO

\section{Motivation}

\subsection{Compositionaility}

Compositionaility is a property of product-level analyses. An analysis is compositional if it is applied to a component of a system rather than the system as a whole. The granularity of that component varies from one analysis to another. In many programming languages, type checkers work at the level of a compilation unit or a module. Intra-procedural analyses on the other hand work at the level of a single procedure. Non-compositional analyses on the other hand include monolithic model checkers that build a state machine of the whole model with no decomposition into smaller units.

\subsection{Modularity}

In the context of this paper modularity is an attribute of Product Line analyses. By modular, we mean that the Product Line analysis uses the corresponding product-level analysis as a sub-module. This can be done in one of two ways:

A product-level analysis can be used as a black box. In such cases the Product Line analysis calls into the product level analysis and aggregates the results for the whole Product Line.

A product-level analysis can be used as a white box. Some automatic technique is used to lift the implementation of the product-level analysis into a Produt Line analysis. Since this process is automated, we still consider this approach modular.

Non-modular Product Line analyses on the other hand are written from scratch independently of the corresponding product level analysis.

\section{Taxonomy}

\subsection{Compositional and Modular}

\subsection{Compositional and Non-Modular}

\subsection{Non-Compositional and Modular}

\subsection{Non-Compositional and Non-Modular}

\section{Discussion}
TODO

\section{Conclusion}
TODO

\bibliographystyle{ieeetr}
\bibliography{taxonomy} 
\end{document}