### 2018-03-07 doc/quals/background.tex
final version of the quals document

\documentclass[11pt]{article}
\begin{document}

My name is Ramy Shahin, a second year PhD student supervised by Marsha Chechik. I had my M.Sc. degree in Computer Science from Michigan State University in 2001, and since then I had worked as a Software Engineer for several companies, including Microsoft, IBM and Amazon.

My research focuses on analysis techniques for Software Product Lines (SPLs). In particular, I have been studying previous attempts at lifting product-level analyses to SPLs. My goal is to automate this kind of lifting (or at least do it systematically).

One attempt I had in this direction is lifting lambda-calculus programs to SPLs. The assumption here is that an analysis is a program written in Typed Lambda Calculus (TLC), and with some type-directed code rewriting rules a lifted version of the analysis can be automatically generated. Examples include type checkers, static analyses and model checkers. The goal is to have an automatically lifted analysis with time complexity polynomial in the SPL size as opposed to naive liftings that have exponential time complexity in the number of features.

Lifting TLC would allow us to lift any program written in it, not just analyses. Program and model transformations in particular can be lifted to SPLs as well. In particular, we are working with General Motors (GM) to see how their Assurance Case analyses and transformations can be lifted to their SPLs.

\end{document}